---
title: "Tidy Tuesday Assignment - Week10"
author: "Derek Leask"
date: "11/23/2021"
output: 
  html_document:
    number_sections: true
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Loading the packages used

##trying out the tmap package for the first time
```{r}
library(tidyverse)
library(janitor)
library(here)
library(viridis)
library(lubridate)
library(tidytuesdayR)

library(tmap)
library(sf)
library(spData)
library(magick)
library(gifski)

```


# Loading the data directly from the tidy tuedsay directory
```{r}

tuesdata <- tidytuesdayR::tt_load(2020, week = 36)

key_crop_yields <- tuesdata$key_crop_yields%>%
  clean_names()

```

# Pivoting the data to a long table format and changed the "code" column to "iso_a3" to match with the world data for joining

```{r}
long_crops <- key_crop_yields %>% 
  pivot_longer(cols = 4:last_col(),
               names_to = "crop", 
               values_to = "crop_production") %>% 
  mutate(crop = str_remove_all(crop, " \\(tonnes per hectare\\)")) %>% 
  set_names(nm = names(.) %>% tolower())
  
long_crops<-mutate(long_crops, iso_a3=code)


long_crops
```

# Ran this code to tell the package i want to plot the GIF and not just to view it
```{r}
tmap_mode("plot")
data(World)
```
# Merged the crop data to the world data i got from the tmap package and converted it into sf format
```{r}
map <- merge(long_crops, World[, c(5, 1)], by = "iso_a3", all.y = TRUE)

map <- st_as_sf(map)
```




# Tried to make a facet map, but there were too many years and countries for the facet to be legible, so opted for the animation

```{r}
tm_shape(map) +
  tm_polygons(
    col = "crop_production",
    style = "cont",
    pal = viridis(10, direction = -1),
    title = "Crop Production"
    ) +
  tm_facets(by = "year", ncol = 4) +
  tm_layout(legend.outside.size = 0.2)
```

# ran the this code to get the style guides and color pallates for tmap, which downloads .jpegs of the color pallets and saves them in the project directory
```{r}
#tmap_style_catalogue(path = "./tmap_style_previews", styles = NA)

#tmap_style_catalog(path = "./tmap_style_previews", styles = NA)
```

# ran the animation, but tmap isn't as intuitive as the ggplot package.  Had issues with adjusting the titles and legends, but was able to export as a gif!


```{r}
animation <- tm_shape(map)+
  tm_polygons(
     col = "crop_production",
      style = "cont",
     pal = viridis(10, direction = -1)) +
  tm_facets(along = "year") +
  tm_layout(legend.position = c("left", "bottom"))
  
  tmap_animation(
    animation, filename = "map.gif",
    delay = 75, width = 1200, height = 600, 
    )
```





